---
title: "1.7.22"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(sf)

HPRM_full <- read_rds("~/HPRM/data/us_df_full.rds")
```

#select Bay Area counties: Alameda, Contra Costa, Marin, Napa, San Francisco, San Mateo, Santa Clara, Solano, and Sonoma 

```{r}
HPRM_bayarea <- HPRM_full %>%
  filter(grepl('Alameda|Contra Costa|Marin|Napa|San Francisco|San Mateo|Santa Clara|Solano|Sonoma', NAM))

#change GEOID to numeric for merging later
str(HPRM_bayarea)
HPRM_bayarea$GEOID <- as.numeric(HPRM_bayarea$GEOID)
```

#reading in and cleaning PLACE data: PLACES__Census_Tract_Data__GIS_Friendly_Format___2021_release
```{r}
#read in CA PLACES data and filter for Bay Area 9 counties
PLACES_MH_Bay <- read_csv("~/HPRM/data/PLACES_GIS_Census_2021_CA.csv") %>%
  filter(grepl('Alameda|Contra Costa|Marin|Napa|San Francisco|San Mateo|Santa Clara|Solano|Sonoma', CountyName))

```
# read in Housing policy and HPSA data
```{r}
HPSAMH_df <- read_csv("~/HPRM/data/Health_Professional_Shortage_Area_Mental_Health.csv") %>%
  filter(grepl('Alameda|Contra Costa|Marin|Napa|San Francisco|San Mateo|Santa Clara|Solano|Sonoma', COUNTY))

install.packages("xlsx")
library("xlsx")
#read in housing policy directory
bay_area_policy_2017 <- read.xlsx("~/HPRM/data/Bay_Area_Metro_Housing_Policy_Directory_2017.xlsx", sheetIndex =2)

#delete first row (empty)
bay_area_policy_2017 <- bay_area_policy_2017[-1,]

#switch rows and columns
bay_area_policy_2017_t <- data.frame(t(bay_area_policy_2017))

#remove row names
rownames(bay_area_policy_2017_t) <- c()


#change column names to first row then delete first row and blank last two rows
names(bay_area_policy_2017_t) <- bay_area_policy_2017_t[1,]

bay_area_policy_2017_t <- bay_area_policy_2017_t[-1,]
bay_area_policy_2017_t <- bay_area_policy_2017_t[-111,]
bay_area_policy_2017_t <- bay_area_policy_2017_t[-110,]

#Label counties
county_list <- bay_area_policy_2017_t[[1]]

county_list[1:15] <- "Alameda"
county_list[16:35] <- "Contra Costa"
county_list[36:47] <- "Marin"
county_list[48:53] <- "Napa"
county_list[54] <- "San Francisco"
county_list[55:76] <- "San Mateo"
county_list[77:91] <- "Santa Clara"
county_list[92:99] <- "Solano"
county_list[100:109] <- "Sonoma"

county_list

bay_area_policy_2017_t[1] <- county_list

#label county column
colnames(bay_area_policy_2017_t)[1] <- "CountyName"
```


#outer merge of HPRM and PLACES data
```{r}

merge_1 <- merge(HPRM_bayarea, PLACES_MH_Bay, by.x = "GEOID", by.y = "LocationName", all = TRUE)
```
#read in census tract/bay area county and city data, merge with policy data, and merge with merge_1
```{r}
library(readxl)

#read in census tract data
census_tracts_alameda <- read_xls("~/HPRM/data/census_tract_numbers.xls")
names(census_tracts_alameda) <- census_tracts_alameda[1,]
census_tracts_alameda <- census_tracts_alameda[-1,]

census_tracts_contracosta <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 2)
names(census_tracts_contracosta) <- census_tracts_contracosta[1,]
census_tracts_contracosta <- census_tracts_contracosta[-1,]

census_tracts_marin <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 3)
names(census_tracts_marin) <- census_tracts_marin[1,]
census_tracts_marin <- census_tracts_marin[-1,]

census_tracts_napa <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 4)
names(census_tracts_napa) <- census_tracts_napa[1,]
census_tracts_napa <- census_tracts_napa[-1,]

census_tracts_sanfrancisco <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 5)
names(census_tracts_sanfrancisco) <- census_tracts_sanfrancisco[1,]
census_tracts_sanfrancisco <- census_tracts_sanfrancisco[-1,]

census_tracts_sanmateo <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 6)
names(census_tracts_sanmateo) <- census_tracts_sanmateo[1,]
census_tracts_sanmateo <- census_tracts_sanmateo[-1,]

census_tracts_santaclara <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 7)
names(census_tracts_santaclara) <- census_tracts_santaclara[1,]
census_tracts_santaclara <- census_tracts_santaclara[-1,]

census_tracts_solano <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 8)
names(census_tracts_solano) <- census_tracts_solano[1,]
census_tracts_solano <- census_tracts_solano[-1,]

census_tracts_sonoma <- read_xls("~/HPRM/data/census_tract_numbers.xls", sheet = 9)
names(census_tracts_sonoma) <- census_tracts_sonoma[1,]
census_tracts_sonoma <- census_tracts_sonoma[-1,]


#hash dictionary

# assign_hash <- Vectorize(assign, vectorize.args = c("x", "value"))
# get_hash <- Vectorize(get, vectorize.args = "x")
# exists_hash <- Vectorize(exists, vectorize.args = "x")
# 
# alameda <- rep("Alameda", 17)
# alameda_tracts <- na.omit(census_tracts_alameda[[1]])
# 
# hash <- new.env(hash = TRUE, parent = emptyenv(), size = 100L)
# 
# alameda_hash <- assign_hash(alameda_tracts, alameda, hash)

# install.packages("hash")
# library(hash)
# 
# h <- hash()
# alameda <- rep("Alameda", 17)
# alameda_tracts <- na.omit(census_tracts_alameda[[1]])
# 
# h[alameda_tracts]<-alameda
# alameda_tracts
# h[["4271"]]
# 
# 
# albany <- rep("Albany", 6)
# albany_tracts <- na.omit(census_tracts_alameda[[2]])
# 
# berkeley <- rep("Berkeley", )
# 
# h[albany_tracts] <- albany



# #test assigning merge_1 rows with city based on census tract
# merge_1_alamedatest <- mutate(merge_1, city = h[merge_1$TRACTCE])

#label each merge_1 census tract with city
merge_1_city <- mutate(merge_1, "city" = case_when(
 merge_1$NAME.x %in% census_tracts_alameda$Alameda ~ "Alameda",
  NAME.x %in% census_tracts_alameda$Albany ~ "Albany",
  NAME.x %in% census_tracts_alameda$Berkeley ~ "Berkeley",
  NAME.x %in% census_tracts_alameda$Dublin ~ "Dublin",
  NAME.x %in% census_tracts_alameda$Emeryville ~ "Emeryville", #
  NAME.x %in% census_tracts_alameda$Fremont ~ "Fremont",
  NAME.x %in% census_tracts_alameda$Hayward ~ "Hayward",
  NAME.x %in% census_tracts_alameda$Livermore ~ "Livermore", #
  NAME.x %in% census_tracts_alameda$Newark ~ "Newark",
  NAME.x %in% census_tracts_alameda$Oakland ~ "Oakland",
  NAME.x %in% census_tracts_alameda$Piedmont ~ "Piedmont",
  NAME.x %in% census_tracts_alameda$Pleasanton ~ "Pleasanton",
  NAME.x %in% census_tracts_alameda$"San Leandro" ~ "San Leandro",
  NAME.x %in% census_tracts_alameda$"Union City" ~ "Union City",
  NAME.x %in% census_tracts_contracosta$Antioch ~ "Antioch",
  NAME.x %in% census_tracts_contracosta$Brentwood ~ "Brentwood", #       
  NAME.x %in% census_tracts_contracosta$Clayton ~ "Clayton", #
  NAME.x %in% census_tracts_contracosta$Concord ~ "Concord",
  NAME.x %in% census_tracts_contracosta$Danville ~ "Danville", #
  NAME.x %in% census_tracts_contracosta$"El Cerrito" ~ "El Cerrito",
  NAME.x %in% census_tracts_contracosta$Hercules ~ "Hercules", #
  NAME.x %in% census_tracts_contracosta$Lafayette ~ "Lafayette",
  NAME.x %in% census_tracts_contracosta$Martinez ~ "Martinez",
  NAME.x %in% census_tracts_contracosta$Moraga ~ "Moraga", #
  NAME.x %in% census_tracts_contracosta$Oakley ~ "Oakley", #
  NAME.x %in% census_tracts_contracosta$Orinda ~ "Orinda", #
  NAME.x %in% census_tracts_contracosta$Pinole ~ "Pinole", #
  NAME.x %in% census_tracts_contracosta$"Pleasant Hill" ~ "Pleasant Hill", 
  NAME.x %in% census_tracts_contracosta$Richmond ~ "Richmond",
  NAME.x %in% census_tracts_contracosta$"San Pablo" ~ "San Pablo", #
  NAME.x %in% census_tracts_contracosta$`San Ramon` ~ "San Ramon",
  NAME.x %in% census_tracts_contracosta$`Walnut Creek` ~ "Walnut Creek", #
  NAME.x %in% census_tracts_marin$"Belvedere" ~ "Belvedere",
  NAME.x %in% census_tracts_marin$"Corte Madera" ~ "Corte Madera", #
  NAME.x %in% census_tracts_marin$Fairfax  ~"Fairfax", #
  NAME.x %in% census_tracts_marin$"Mill Valley" ~ "Mill Valley", #
  NAME.x %in% census_tracts_marin$Novato ~ "Novato",
  NAME.x %in% census_tracts_marin$Ross ~ "Ross",
  NAME.x %in% census_tracts_marin$"San Anselmo" ~ "San Anselmo", 
  NAME.x %in% census_tracts_marin$"San Rafael" ~ "San Rafael",
  NAME.x %in% census_tracts_marin$Sausalito  ~"Sausalito",#
  NAME.x %in% census_tracts_marin$Tiburon ~ "Tiburon", #
  NAME.x %in% census_tracts_napa$"American Canyon" ~ "American Canyon", #
  NAME.x %in% census_tracts_napa$Calistoga ~ "Calistoga", #
  NAME.x %in% census_tracts_napa$"St. Helena" ~ "St. Helena", #
  NAME.x %in% census_tracts_napa$Yountville ~ "Yountville", #
  NAME.x %in% census_tracts_sanfrancisco$"San Francisco" ~ "San Francisco",
  NAME.x %in% census_tracts_sanmateo$"Half Moon Bay" ~ "Half Moon Bay", #
  NAME.x %in% census_tracts_sanmateo$"Hillsborough" ~ "Hillsborough", #
  NAME.x %in% census_tracts_sanmateo$"Menlo Park" ~ "Menlo Park",
  NAME.x %in% census_tracts_sanmateo$"Millbrae" ~ "Millbrae",
  NAME.x %in% census_tracts_sanmateo$"Pacifica" ~ "Pacifica",
  NAME.x %in% census_tracts_sanmateo$"Portola Valley" ~ "Portola Valley", #
  NAME.x %in% census_tracts_sanmateo$"Redwood City" ~ "Redwood City",
  NAME.x %in% census_tracts_sanmateo$"San Bruno" ~ "San Bruno",
  NAME.x %in% census_tracts_sanmateo$"San Carlos" ~ "San Carlos",
  NAME.x %in% census_tracts_sanmateo$"San Mateo" ~ "San Mateo",
  NAME.x %in% census_tracts_sanmateo$"South San Francisco" ~ "South San Francisco",
  NAME.x %in% census_tracts_sanmateo$"Woodside" ~ "Woodside",
  NAME.x %in% census_tracts_santaclara$"Campbell" ~ "Campbell", #
  NAME.x %in% census_tracts_santaclara$"Cupertino" ~ "Cupertino", #
  NAME.x %in% census_tracts_santaclara$Gilroy ~ "Gilroy", #
  NAME.x %in% census_tracts_santaclara$"Los Altos" ~ "Los Altos",
  NAME.x %in% census_tracts_santaclara$"Los Altos Hills" ~ "Los Altos Hills",#
  NAME.x %in% census_tracts_santaclara$"Los Gatos" ~ "Los Gatos", #
  NAME.x %in% census_tracts_santaclara$"Milpitas" ~ "Milpitas",
  NAME.x %in% census_tracts_santaclara$"Monte Sereno" ~ "Monte Sereno", #
  NAME.x %in% census_tracts_santaclara$"Morgan Hill" ~ "Morgan Hill", #
  NAME.x %in% census_tracts_santaclara$"Mountain View" ~ "Mountain View",
  NAME.x %in% census_tracts_santaclara$"Palo Alto" ~ "Palo Alto",
  NAME.x %in% census_tracts_santaclara$"San Jose" ~ "San Jose",
  NAME.x %in% census_tracts_santaclara$"Santa Clara" ~ "Santa Clara",
  NAME.x %in% census_tracts_santaclara$"Saratoga" ~ "Saratoga",
  NAME.x %in% census_tracts_santaclara$"Sunnyvale" ~ "Sunnyvale",
  NAME.x %in% census_tracts_solano$"Benicia" ~ "Benicia",#
  NAME.x %in% census_tracts_solano$"Dixon" ~ "Dixon",#
  NAME.x %in% census_tracts_solano$"Fairfield" ~ "Fairfield",#
  NAME.x %in% census_tracts_solano$"Rio Vista" ~ "Rio Vista", #
  NAME.x %in% census_tracts_solano$"Suisun City" ~ "Suisun City",#
  NAME.x %in% census_tracts_solano$"Vacaville" ~ "Vacaville",#
  NAME.x %in% census_tracts_solano$"Vallejo" ~ "Vallejo", #
  NAME.x %in% census_tracts_sonoma$"Cloverdale" ~ "Cloverdale",#
  NAME.x %in% census_tracts_sonoma$"Cotati" ~ "Cotati",#
  NAME.x %in% census_tracts_sonoma$"Healdsburg" ~ "Healdsburg", #
  NAME.x %in% census_tracts_sonoma$"Petaluma" ~ "Petaluma", #
  NAME.x %in% census_tracts_sonoma$"Rohnert Park" ~ "Rohnert Park",#
  NAME.x %in% census_tracts_sonoma$"Santa Rosa" ~ "Santa Rosa", #
  NAME.x %in% census_tracts_sonoma$"Sebastopol" ~ "Sebastopol", #
  NAME.x %in% census_tracts_sonoma$"Sonoma" ~ "Sonoma", #
  NAME.x %in% census_tracts_sonoma$"Windsor" ~ "Windsor",#
TRUE ~ "NA"))

test2<-merge_1_city %>%
group_by(city) %>%
summarize(count = n())

#outer merge of merge_1 and policy data

# "4271" %in% census_tracts_alameda$Alameda

#merge_2 <- merge(merge_1, bay_area_policy_2017_t, by = "city", all = TRUE)

#outer merge of merge_2 and HPSA data
#merge_2 <- merge(merge_1_city, HPSAMH_df, by = , all = TRUE)

```


#visualizing data
```{r MENTAL HEALTH}

merge_1MH <- merge_1 %>%
  filter(MeasureId == "MHLTH")

library(dplyr)
library(knitr)
library(ggplot2)
table(merge_1MH$NeighType)
ggplot(merge_1, aes(x=pBlack))+
  geom_histogram()
#percent race vs prevalence MH outcome

MHsum <- merge_1MH %>%
  group_by(NeighType) %>%
  summarize("Data_Value" = mean(Data_Value))

ggplot(MHsum, aes(x=NeighType, y= Data_Value)) +
  geom_bar(stat="identity")+
	      theme(axis.text.x = element_text(angle = -45, hjust = 0))+ 
        labs(title="Neighborhood Type vs. Average prevalence of MH Outcome", y="Average prevalence MH Outcome", x="Neighborhood type", caption="MH Outcome = presence of poor mental health for more than 14 days out of 30 days. Data from CDC PLACES Survey 2019")


#average prevalence per tract in that NeighType
ggplot(merge_1MH, aes(x=pBlack, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()

ggplot(merge_1MH, aes(x=pWhite, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()
ggplot(merge_1MH, aes(x=pLatinx, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()

ggplot(merge_1MH, aes(x=pAsian, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()


#percent race and HPRM scale
ggplot(merge_1MH, aes(x=pLatinx, y=hprm_scale)) +
  geom_point(stat="identity") +
  geom_smooth()
ggplot(merge_1MH, aes(x=pBlack, y=hprm_scale)) +
  geom_point(stat="identity") +
  geom_smooth()

#HRPM scale and MH outcome prevalence
ggplot(merge_1MH, aes(x=hprm_scale, y=Data_Value)) +
  geom_point(stat="identity")

M1 <- lm(data = merge_1MH, Data_Value ~ hprm_scale)
summary(M1)

M2 <- lm(data = merge_1MH, Data_Value ~ hprm_scale + pBlack + pLatinx + pAsian + pWhite)
summary(M2)

M3 <- lm(data = merge_1MH, Data_Value ~ hprm_scale + pBlack + pLatinx + pAsian)
summary(M3)
```

```{r DEPRESSION}
merge_1D <- merge_1 %>%
  filter(MeasureId == "DEPRESSION")

library(dplyr)
library(knitr)
library(ggplot2)
table(merge_1D$NeighType)
ggplot(merge_1D, aes(x=pBlack))+
  geom_histogram()
#percent race vs prevalence Depression outcome

DEPsum <- merge_1D %>%
  group_by(NeighType) %>%
  summarize("Data_Value" = mean(Data_Value))

ggplot(DEPsum, aes(x=NeighType, y= Data_Value)) +
  geom_bar(stat="identity")+
	      theme(axis.text.x = element_text(angle = -45, hjust = 0))+ 
        labs(title="Neighborhood Type vs. Average prevalence of Depression Outcome", y="Average prevalence Depression", x="Neighborhood type", caption="Depression Outcome = presence of diagnosis of depression (including depression, major depression, dysthymia, or minor depression) among respondents age 18 or older. Data from CDC PLACES Survey 2019")

#average prevalence per tract in that NeighType
ggplot(merge_1D, aes(x=pBlack, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()

ggplot(merge_1D, aes(x=pWhite, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()
ggplot(merge_1D, aes(x=pLatinx, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()

ggplot(merge_1D, aes(x=pAsian, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()


#percent race and HPRM scale
ggplot(merge_1D, aes(x=pLatinx, y=hprm_scale)) +
  geom_point(stat="identity") +
  geom_smooth()
ggplot(merge_1D, aes(x=pBlack, y=hprm_scale)) +
  geom_point(stat="identity") +
  geom_smooth()

#HRPM scale and MH outcome prevalence
ggplot(merge_1D, aes(x=hprm_scale, y=Data_Value)) +
  geom_point(stat="identity")

M1 <- lm(data = merge_1D, Data_Value ~ hprm_scale)
summary(M1)

M2 <- lm(data = merge_1D, Data_Value ~ hprm_scale + pBlack + pLatinx + pAsian + pWhite)
summary(M2)

M3 <- lm(data = merge_1D, Data_Value ~ hprm_scale + pBlack + pLatinx + pAsian)
summary(M3)
```

```{r SLEEP}
merge_1S <- merge_1 %>%
  filter(MeasureId == "SLEEP")

library(dplyr)
library(knitr)
library(ggplot2)
table(merge_1MH$NeighType)
ggplot(merge_1, aes(x=pBlack))+
  geom_histogram()
#percent race vs prevalence sleep outcome

SLEEPsum <- merge_1S %>%
  group_by(NeighType) %>%
  summarize("Data_Value" = mean(Data_Value))

ggplot(SLEEPsum, aes(x=NeighType, y= Data_Value)) +
  geom_bar(stat="identity")+
	      theme(axis.text.x = element_text(angle = -45, hjust = 0))+ 
        labs(title="Neighborhood Type vs. Average prevalence of Sleep Outcome", y="Average prevalence MH Outcome", x="Neighborhood type", caption="Sleep Outcome = usually sleeping less than 7 hours per night among respondents age 18 or older. Data from CDC PLACES Survey 2019")

#average prevalence per tract in that NeighType
ggplot(merge_1S, aes(x=pBlack, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()

ggplot(merge_1S, aes(x=pWhite, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()
ggplot(merge_1S, aes(x=pLatinx, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()

ggplot(merge_1S, aes(x=pAsian, y=Data_Value)) +
  geom_point(stat="identity") +
  geom_smooth()


#percent race and HPRM scale
ggplot(merge_1S, aes(x=pLatinx, y=hprm_scale)) +
  geom_point(stat="identity") +
  geom_smooth()
ggplot(merge_1S, aes(x=pBlack, y=hprm_scale)) +
  geom_point(stat="identity") +
  geom_smooth()

#HRPM scale and MH outcome prevalence
ggplot(merge_1S, aes(x=hprm_scale, y=Data_Value)) +
  geom_point(stat="identity")

M1 <- lm(data = merge_1S, Data_Value ~ hprm_scale)
summary(M1)

M2 <- lm(data = merge_1S, Data_Value ~ hprm_scale + pBlack + pLatinx + pAsian + pWhite)
summary(M2)

M3 <- lm(data = merge_1S, Data_Value ~ hprm_scale + pBlack + pLatinx + pAsian)
summary(M3)
```